A introduction to lattice trapdoor

### 2023-11-24 co687/project/03-nearest-plane-cvp.tex
Some new formatting

1. If $B$ is an empty set of basis, return $\mathbf{0}$
2. $B^\ast \leftarrow \text{GramSchmidt}(B)$
3. $c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil$, where $\mathbf{b}_n^\ast$ is the last base vector in $B^\ast$
4. $B' = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}]$ is obtained by removing the last base vector from $B$
5. return $c\mathbf{b}_n + \text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$

>>>>

\begin{enumerate}
    \item If $B$ is an empty set of basis, return $\mathbf{0}$
    \item $B^\ast \leftarrow \text{GramSchmidt}(B)$
    \item $c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil$, where $\mathbf{b}_n^\ast$ is the last base vector in $B^\ast$
    \item $B' = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}]$ is obtained by removing the last base vector from $B$
    \item return $c\mathbf{b}_n + \text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$
\end{enumerate}

### 2023-11-24 co687/project/04-ggh-trapdoor.tex
Some new formatting

One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper *"Public-key cryptosystem from lattice reduction problem"*. At a high level, the trapdoor is parameterized by three items: a "bad" basis $B$, a "good" basis $R$, and an error bound $\sigma$. In the forward direction, the function maps a pair of lattice coordinate $\mathbf{v} \in \mathbb{Z}^n$ and a small error vector $\mathbf{e} \leftarrow \{ -\sigma, \sigma \}^n$ to $\mathbf{x} = B\mathbf{v} + \mathbf{e} \in \mathbb{R}^n$. If the parameters are generated correctly, then the closest lattice point in $\mathcal{L}(B)$ is exactly $B\mathbf{v}$.

@@

1. Parameter generation  
The main security parameter in this scheme is the number of dimensions $n$ of the lattice. In the original paper, the authors claimed $n \approx 150$ is sufficient for making inverting the function without the private basis hard and $n \approx 250$ should be a safe choice for the foreseeable future.
2. Key generation  
First generate the "good" basis $R \in \mathbb{R}^{n \times }$, then apply some (unimodular) transformation to $R$ to obtain the "bad" basis $B$. The error bound $\sigma > 0 \in \mathbb{R}$ is dependent on the choice of $R$ and the choice of "probability of inversion error" $\epsilon >= 0$, which will be discussed in a later section.
3. Forward evaluation  
$f_{B}(\mathbf{v}, \mathbf{e}) = B\mathbf{v} + \mathbf{e}$, where $\mathbf{v} \in \{-n, \ldots, n\}^n$ and $\mathbf{e} \in \{-\sigma, \sigma\}$. According to the authors, the choice of bounds for values of $\mathbf{v}$ is arbitrary and not a significant contributor to the overall security of the scheme.
4. Inversion  
Denote the output by $\mathbf{x} = f_{B}(\mathbf{v}, \mathbf{e})$, first attempt to recover the integer coordinate $\mathbf{v} \leftarrow B^{-1}R\lfloor R^{-1}\mathbf{x} \rceil$. From here it is trivial to recompute the lattice point $B\mathbf{v}$ and recover the error term $\mathbf{e} = \mathbf{x} - B\mathbf{v}$.

>>>>

One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper \textit{"Public-key cryptosystem from lattice reduction problem"}. At a high level, the trapdoor is parameterized by three items: a "bad" basis $B$, a "good" basis $R$, and an error bound $\sigma$. In the forward direction, the function maps a pair of lattice coordinate $\mathbf{v} \in \mathbb{Z}^n$ and a small error vector $\mathbf{e} \leftarrow \{ -\sigma, \sigma \}^n$ to $\mathbf{x} = B\mathbf{v} + \mathbf{e} \in \mathbb{R}^n$. If the parameters are generated correctly, then the closest lattice point in $\mathcal{L}(B)$ is exactly $B\mathbf{v}$.

@@

\begin{enumerate}
    \item \textbf{Parameter generation} The main security parameter in this scheme is the number of dimensions $n$ of the lattice. In the original paper, the authors claimed $n \approx 150$ is sufficient for making inverting the function without the private basis hard and $n \approx 250$ should be a safe choice for the foreseeable future.
    \item \textbf{Key generation} First generate the "good" basis $R \in \mathbb{R}^{n \times }$, then apply some (unimodular) transformation to $R$ to obtain the "bad" basis $B$. The error bound $\sigma > 0 \in \mathbb{R}$ is dependent on the choice of $R$ and the choice of "probability of inversion error" $\epsilon >= 0$, which will be discussed in a later section.
    \item \textbf{Forward evaluation} $f_{B}(\mathbf{v}, \mathbf{e}) = B\mathbf{v} + \mathbf{e}$, where $\mathbf{v} \in \{-n, \ldots, n\}^n$ and $\mathbf{e} \in \{-\sigma, \sigma\}$. According to the authors, the choice of bounds for values of $\mathbf{v}$ is arbitrary and not a significant contributor to the overall security of the scheme.
    \item \textbf{Inversion} Denote the output by $\mathbf{x} = f_{B}(\mathbf{v}, \mathbf{e})$, first attempt to recover the integer coordinate $\mathbf{v} \leftarrow B^{-1}R\lfloor R^{-1}\mathbf{x} \rceil$. From here it is trivial to recompute the lattice point $B\mathbf{v}$ and recover the error term $\mathbf{e} = \mathbf{x} - B\mathbf{v}$.
\end{enumerate}

### 2023-11-24 co687/project/main.tex
Some new formatting

\documentclass[letterpaper,12pt]{article}

\usepackage{amsmath,amssymb,amstext}
\setlength{\marginparwidth}{0pt} % width of margin notes
% N.B. If margin notes are used, you must adjust \textwidth, \marginparwidth
% and \marginparsep so that the space left between the margin notes and page
% edge is less than 15 mm (0.6 in.)
\setlength{\marginparsep}{0pt} % width of space between body text and margin notes
\setlength{\evensidemargin}{0.125in} % Adds 1/8 in. to binding side of all 
% even-numbered pages when the "twoside" printing option is selected
\setlength{\oddsidemargin}{0.125in} % Adds 1/8 in. to the left of all pages when "oneside" printing is selected, and to the left of all odd-numbered pages when "twoside" printing is selected
\setlength{\textwidth}{6.375in} % assuming US letter paper (8.5 in. x 11 in.) and side margins as above

\begin{document}

>>>>

\documentclass[letterpaper,12pt]{article}
\usepackage{amsmath,amssymb,amstext}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amssymb}

\title{A introduction to lattice trapdoor}
\author{Ganyu Xu (g66xu)}
\date{Fall, 2023}

\begin{document}
\maketitle
